%!TEX root = 0_QE_Report.tex

\begin{acknowledgements}

I would firstly likely to thank Dr. Michael Lees for giving me the opportunity and the inspiration to pursue my dreams. Right from the days of co-supervising my Final Year Project, he has always been a reliable and uncompromising critic who's been there to help, advise and discuss about everything ranging from my research to gadgets to football. The weekly team meetings over coffee and his emphasis on encouraging everyone in his team to discuss their projects with each other has not only helped me develop a lot of my ideas but has made PhD life a lot more interesting when things get boring; and bearable when the going gets tough. 

I would also like to thank Heiko Aydt and Hu Nan for being sounding boards for my ideas and friends in an environment where I've sometimes felt lost. I'm also grateful to them for letting me be a part of some of the really interesting work they're doing. Their experiences in working on a PhD at NTU has helped me avoid and handle all the major and minor hurdles that I have faced as a student so far.

I would also like to express my sincere gratitude to the School of Computer Engineering, NTU and the Parallel and Distributed Computing Center for providing me with the opportunity, the funds and, above all, the excellent facilities to pursue a PhD. A special thanks is also due to Kenneth (Poliran Kenneth Caballes) and Irene (Ng-Goh Siew Lai) for making working at PDCC so easy and convenient.

Finally I would like to thank my family for their unflinching support and belief in me; and my girlfriend, Aparna, for listening to my constant complaints and rants, and never failing to cheer me up. \\
\\
\hspace{0in}
Vaisagh Viswanathan T\\
Dec 2011\\
{Nanyang Technological University, Singapore}\\
\end{acknowledgements}